---
title: "Direct Care Financial Report"
date: "`r format(params$generated_on, '%B %d, %Y')`"
output:
  html_document:
    theme: cosmo
    toc: true
    toc_float:
      collapsed: false
    css: report_style.css
    self_contained: true
params:
  practice_name:   "Practice"
  practice_id:     ""
  forecast_type:   "breakeven"
  result:          NULL
  overhead_monthly: NULL
  generated_on:    !r Sys.Date()
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 9, fig.height = 4.5, dpi = 150)
library(ggplot2)
library(dplyr)
library(scales)
library(directCareAnalytics)   # plot + interpret helpers
```

---

<div class="rds-header">
  <div class="rds-brand">Raborn Decision Sciences</div>
  <div class="rds-tagline">Financial Analytics for Direct Care Practices</div>
</div>

---

# `r params$practice_name`

**Practice ID:** `r params$practice_id`
**Report type:** `r tools::toTitleCase(params$forecast_type)`
**Generated:** `r format(params$generated_on, "%B %d, %Y")`

---

## Practice Overview

```{r overhead-summary}
om <- params$overhead_monthly
if (!is.null(om) && nrow(om) > 0) {
  months_of_data <- nrow(om)
  avg_overhead   <- mean(om$total_overhead, na.rm = TRUE)
  total_refunds  <- sum(om$total_refunds, na.rm = TRUE)

  cat(
    "- **Months of overhead data:** ", months_of_data, "\n",
    "- **Average monthly overhead:** ", dollar(avg_overhead), "\n",
    "- **Total refunds / credits recorded:** ", dollar(total_refunds), "\n",
    sep = ""
  )
}
```

```{r overhead-chart, fig.cap="Monthly overhead by period"}
om <- params$overhead_monthly
if (!is.null(om) && nrow(om) > 0) {
  om |>
    mutate(period = as.Date(paste(year, month, "01", sep = "-"))) |>
    ggplot(aes(x = period, y = total_overhead)) +
    geom_col(fill = "#1e3a5f", alpha = 0.85) +
    geom_line(aes(y = gross_overhead), colour = "#c0392b",
              linewidth = 0.8, linetype = "dashed") +
    scale_y_continuous(labels = dollar_format()) +
    scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") +
    labs(
      title    = "Monthly Overhead",
      subtitle = "Bars = net overhead  |  Dashed line = gross overhead (before refunds)",
      x = NULL, y = "Amount ($)"
    ) +
    theme_minimal(base_size = 12) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```

---

## Forecast Results

```{r forecast-plot, fig.cap="Forecast output"}
res  <- params$result
type <- params$forecast_type

if (!is.null(res)) {
  if (type == "breakeven") {
    directCareAnalytics:::plot_forecast_breakeven(res)
  } else if (type == "revenue") {
    directCareAnalytics:::plot_forecast_revenue(res)
  } else if (type == "target") {
    directCareAnalytics:::plot_forecast_target(res)
  }
}
```

```{r key-metrics}
res  <- params$result
type <- params$forecast_type
if (!is.null(res)) {
  if (type == "breakeven") {
    cat(
      "| Metric | Value |\n",
      "|--------|-------|\n",
      "| Break-even date |",
        if (is.na(res$breakeven_date)) "Not reached in horizon"
        else format(res$breakeven_date, "%B %Y"), "|\n",
      "| Months to break-even |",
        if (is.na(res$periods_to_breakeven)) "—"
        else res$periods_to_breakeven, "|\n",
      "| Current surplus / deficit |",
        dollar(res$current_surplus_deficit), "|\n",
      sep = ""
    )
  } else if (type == "revenue") {
    cat(
      "| Metric | Value |\n",
      "|--------|-------|\n",
      "| Current monthly revenue |", dollar(res$current_revenue), "|\n",
      "| Projected (end of horizon) |",
        dollar(tail(res$forecast_data$revenue_forecast, 1)), "|\n",
      "| Forecast method |", res$method, "|\n",
      sep = ""
    )
  } else if (type == "target") {
    cat(
      "| Metric | Value |\n",
      "|--------|-------|\n",
      "| Target date |",
        if (is.na(res$target_date)) "Not reached in horizon"
        else format(res$target_date, "%B %Y"), "|\n",
      "| Revenue required now |", dollar(res$required_revenue_now), "|\n",
      "| Current gap |", dollar(res$current_gap), "|\n",
      sep = ""
    )
  }
}
```

---

## Interpretation

```{r interpretation, results='asis'}
res  <- params$result
type <- params$forecast_type
name <- params$practice_name

if (!is.null(res)) {
  txt <- if (type == "breakeven") {
    directCareAnalytics:::interpret_breakeven(res, name)
  } else if (type == "revenue") {
    directCareAnalytics:::interpret_revenue(res, name)
  } else if (type == "target") {
    directCareAnalytics:::interpret_target(res, name)
  } else ""
  cat(txt)
}
```

---

<div class="rds-footer">
  Prepared by **Raborn Decision Sciences** &nbsp;·&nbsp;
  `r format(params$generated_on, "%B %Y")` &nbsp;·&nbsp;
  Confidential — for internal practice use only
</div>
